Report for the Deep Learning Course Assignment 2

### 2016-12-05 practical_3/report/report_methenitis.tex
update

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{fancyvrb,xcolor}
\usepackage{graphicx}

@@

\author{
  Georgios Methenitis \\
  \texttt{georgios.methenitis@cwi.nl}
}

@@

\begin{abstract}
In this assignment I got familiar with TensorFlow. I created a single hidden layer network with ReLU activation function to learn features for the CIFAR10 dataset.
\end{abstract}

\section{Task 1}
\begin{enumerate}
\item TensorFlow \emph{variables} are stored structures containing tensors, \emph{constants} are constant variables in Tensorflow, \emph{placeholders} are used in the place of variables at graph initialization when there are no variables. The differences are that constants cannot change their values in contrast with variables, and that placeholder are replacing variables at initialization phase.
\item There are two ways to declare variable in TensorFlow:
\begin{itemize}
\item \begin{small}
\begin{verbatim}
var = tf.get_variable("var_name", shape=[..., ...])
\end{verbatim}
\end{small}
\item \begin{small}
\begin{verbatim}
var = tf.Variable(tf.zeros([..., ...]), name="var_name")
\end{verbatim}
\end{small}
\end{itemize}

\item The command \texttt{tf.shape(x)} returns the shape of the input of tensor \texttt{x} in an 1-D vector, while \texttt{x.get\_shape()} return the size of the tensor.
\item
\item
\item A computational \emph{graph}, \emph{tensors} for the variables and \emph{operator}.
\item In variable scopes, variables can be created, adopting the prefix of the scope in their name, variable scopes enable variable sharing in different places in the code. The differences between name scopes and variable scopes is in the former creates an hierarchy for operations and variables in the graph, while the latter helps mostly in variable sharing.
\item You can freeze a variable tensor during training by setting \texttt{trainable=True} upon creating the variable, this way it will not get updated during optimization.
\item
\item
\end{enumerate}

\section{Task 2}
In this task I implemented the following functions in \texttt{mlp.py}, more specifically the \texttt{inference}, \texttt{loss}, and \texttt{accuracy}.

\begin{small}
\begin{verbatim}
  def inference(self, x):
    with tf.variable_scope('hidden', reuse=None):
      W = tf.get_variable("weights", shape=[x.get_shape()[1], self.n_hidden[0]],
       initializer=self.weight_initializer, regularizer=self.weight_regularizer)
      tf.histogram_summary("hidden_weights", W)
      b = tf.Variable(tf.zeros([self.n_hidden[0]]), name="bias")
      tf.histogram_summary("hidden_bias", b)

    with tf.variable_scope('output', reuse=None):
      w_out = tf.get_variable("weights", shape=[self.n_hidden[0], self.n_classes],
       initializer=self.weight_initializer, regularizer=self.weight_regularizer)
      tf.histogram_summary("output_weights", w_out)
      b_out = tf.Variable(tf.zeros([self.n_classes]), name="bias")
      tf.histogram_summary("output_bias", b_out)

    with tf.name_scope('relu_layer'):
      input = self.activation_fn(tf.matmul(x, W) + b)
    with tf.name_scope('linear_layer'):
      output = tf.matmul(input, w_out) + b_out

    return output

  def loss(self, logits, labels):

    loss = tf.reduce_mean(tf.nn.softmax_cross_entropy_with_logits(logits, labels, name=None))
    with tf.variable_scope("hidden", reuse=True):
      loss += 0.5 * tf.reduce_sum(tf.get_variable("weights")**2)
    with tf.variable_scope("output", reuse=True):
      loss += 0.5 * tf.reduce_sum(tf.get_variable("weights")**2)

    return loss

  def accuracy(self, logits, labels):
    correct_prediction = tf.equal(tf.argmax(labels, 1), tf.argmax(logits, 1))
    accuracy = tf.reduce_mean(tf.cast(correct_prediction, tf.float32))
    return accuracy
\end{verbatim}
\end{small}
I also implemented the function \texttt{train} in the file \texttt{train\_mlp.py}:
\begin{small}
\begin{verbatim}
def train():
  # placeholder for input and output tensors allowing to create operations and the computation graph
  x = tf.placeholder(tf.float32, shape=[None, 3072])
  y_ = tf.placeholder(tf.float32, shape=[None, 10])

  # model, namescopes for summary
  model = MLP()
  x_ = model.inference(x)
  with tf.name_scope('loss'):
    loss = model.loss(x_, y_)
  with tf.name_scope('accuracy'):
    accuracy = model.accuracy(x_, y_)
  with tf.name_scope('train_step'):
    train_step = tf.train.GradientDescentOptimizer(LEARNING_RATE_DEFAULT).minimize(loss)

  # session
  init = tf.initialize_all_variables()
  sess = tf.Session()
  sess.run(init)

  # merged summary for the graph, the loss and accuracy
  tf.scalar_summary('accuracy', accuracy)
  tf.scalar_summary('loss', loss)
  tf.histogram_summary("logits", x_) # logits histogram
  merged = tf.merge_all_summaries()

  train_writer = tf.train.SummaryWriter(LOG_DIR_DEFAULT + '/train', sess.graph)
  test_writer = tf.train.SummaryWriter(LOG_DIR_DEFAULT + '/test', sess.graph)

  for i in range(1, MAX_STEPS_DEFAULT+1):
      batch_xs, batch_ys = cifar10.train.next_batch(BATCH_SIZE_DEFAULT)
      __, Summary, l, acc = sess.run([train_step, merged, loss, accuracy],
      feed_dict={x: batch_xs, y_: batch_ys})
      train_writer.add_summary(Summary, i)

      if i % 100 == 0.0:
          batch_xs, batch_ys = cifar10.test.images, cifar10.test.labels
          __, Summary, l, acc = sess.run([train_step, merged, loss, accuracy],
           feed_dict={x: batch_xs, y_: batch_ys})
          test_writer.add_summary(Summary, i)

  train_writer.close()
\end{verbatim}
\end{small}

Figure~\ref{fig:acclosstrain} presents the accuracy and the loss in the training set recorded for every epoch.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{loss.png}\	
\includegraphics[width=0.45\textwidth]{accuracy.png}
\caption{Recorded accuracy and loss for the training set.}
\label{fig:acclosstrain}
\end{figure}

Figure~\ref{fig:acclosstest} presents the accuracy and the loss in the test set recorded for every 100-epochs.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{accuracy_t.png}\	
\includegraphics[width=0.45\textwidth]{loss_t.png}
\caption{Recorded accuracy and loss for the test set.}
\label{fig:acclosstest}
\end{figure}

Figure~\ref{fig:graph} presents the graph of the created model as this is illustrated by TensorBoard.

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{graph.png}
\caption{Graph of the model.}
\label{fig:graph}
\end{figure}

Figure~\ref{fig:histogramsunits} presents the histograms of the weights and the biases in every unit of the model, in the hidden and the output units respectively.

\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{hidden_weights.png}
\includegraphics[width=0.3\textwidth]{hidden_bias.png}
\includegraphics[width=0.3\textwidth]{output_weights.png}
\includegraphics[width=0.3\textwidth]{output_bias.png}
\caption{Histograms of weights and bias in the hidden layer, and weights and bias in the output linear layer.}
\label{fig:histogramsunits}
\end{figure}

Figure~\ref{fig:histogramlogits} presents the histogram of the logits.

\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{logits.png}
\caption{Histograms of logits.}
\label{fig:histogramlogits}
\end{figure}

\section{Conclusion}
Should contain conclusion of this study.
For example, you can try to answer the following questions.
What was done during this assignment? What features of TensorBoard were positive and what were negative for implementing MLP model and performing the experiments?
What are the main insights you got from the study of the MLP model on CIFAR10 dataset?

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}

@@

\author{
  David S.~Hippocampus \\
  \texttt{hippo@cs.cranberry-lemon.edu}
}

@@

\begin{abstract}
Should contain information about the current task and the summary of the study of the CNN models on CIFAR10 dataset.

\end{abstract}

\section{Task 1}
Should contain all needed information about Task 1 and report of all your experiments for that task.

Figure~\ref{fig:1}, \ref{fig:2}

\begin{figure}[h!]
\centering
\includegraphics[height=3.2cm]{acc1.png}\	
\includegraphics[height=3.2cm]{loss1.png}
\caption{Accuracy (left) and loss on the training set with batch size of $128$ samples. The accuracy approximates $1.0$ in the training set implying that the model overfits the training set.}
\label{fig:1}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[height=3.2cm]{acc2.png}\	
\includegraphics[height=3.2cm]{loss2.png}
\caption{Accuracy (left) and loss on the test set with batch size of $128$ samples. The accuracy is $0.88$ after $15000$ steps.}
\label{fig:2}
\end{figure}

\subsection{Smaller batch size}

Figure~\ref{fig:3}, \ref{fig:4}

\begin{figure}[h!]
\centering
\includegraphics[height=3.2cm]{acc3.png}\	
\includegraphics[height=3.2cm]{loss3.png}
\caption{Accuracy (left) and loss on the training set with batch size of $128$ samples. The accuracy approximates $1.0$ in the training set implying that the model overfits the training set.}
\label{fig:3}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[height=3.2cm]{acc4.png}\	
\includegraphics[height=3.2cm]{loss4.png}
\caption{Accuracy (left) and loss on the test set with batch size of $128$ samples. The accuracy is $0.88$ after $15000$ steps.}
\label{fig:4}
\end{figure}

\section{Task 2}
Should contain all needed information about Task 2 and report of all your experiments for that task.

\section{Task 3}
Should contain all needed information about Task 3 and report of all your experiments for that task.

\section{Conclusion}
Should contain conclusion of this study.

%\section*{References}
%
%\small
%
%[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms
%for connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and
%T.K.\ Leen (eds.), {\it Advances in Neural Information Processing
%  Systems 7}, pp.\ 609--616. Cambridge, MA: MIT Press.
%
%[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS:
%  Exploring Realistic Neural Models with the GEneral NEural SImulation
%  System.}  New York: TELOS/Springer--Verlag.
%
%[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of
%learning and recall at excitatory recurrent synapses and cholinergic
%modulation in rat hippocampal region CA3. {\it Journal of
%  Neuroscience} {\bf 15}(7):5249-5262.
